A Computational Account of Default-Mode Function\\ by Control Theory and Reinforcement Learning

### 2016-10-02 paper/paper.tex
rushworth2016
truc

The dmPFC subserves predominantly ambiguous amodal processes
across time, space, and content domains in
sensory-independent top-down pathways.
This DMN node has been described as a “mental sketchpad” (Goldman-Rakic, 1996),
potentially implicated in de-novo generation and binding
of meaning representations instructed by stored semantics and memories.
%
The dmPFC may thus enable inference, representation, and assessment
of one's own and other individuals' action and thoughts.
Patients with neurological lesions in the prefrontal cortex
are known to adapt to novel situations and stimuli with difficulty
(Stuss, D. T. \& Benson, D. F. The Frontal Lobes, Raven, New York, 1986).
Specifically, neural activity in human dmPFC
reflected expectations about other peoples actions and errors thereof
(Suzuki et al., 2012; van Schie et al., 2004).
dmPFC activity indeed explained the proficiency decline
of inferring other peoples thoughts in the elderly (Moran et al., 2012).
Some dmPFC neurons in macaque monkeys exhibited a preference
for processing others', rather than own, behavior
with fine-grained adjustment of contextual circumstances (Yoshida et al., 2011).
%
Such highly abstract neural computations necessarily rely on the
generation of probabilistic internal information drawing from
episodic memory retrieval, generative construction processes,
and explicit knowledge of the external world.
%
According to computational neuroimaging experiments,
the dmPFC activity preferentially models stimulus-value associations of
possible actions that are not actually executed,
whereas the vmPFC activity models value in the environment that leads performed actions
{Nicolle, 2012 \#3750}.

@@

where the expectation is over all possible trajectories successor
trajectories of state $s_t$.

By construction, in general, maximizing the expectation in  \eqref{eq:qval} is intractable due to
the enormous number of possible trajectories to consider.
One popular simplification that removes this issue is to only consider the sub-family of
\textit{deterministic policies}, which map each state to the best single best action which
should be taken at thin state, and therefore define functions from states to actions.

@@

\paragraph{Free Energy Principle.}
\begin{itemize}
  \item Ortega et al. 2013 \cite{ortega2013thermodynamics} ``Thermodynamics as a theory of decision-making with information-processing costs''
\item Friston's critic of RL \cite{fristonAIorRL} (he proposes "active inference''): "This equation (equation 18) comes from the theory of dynamic programming,
pioneered in the 1950s by Richard Bellman and colleagues [2]. To
optimise control a~pð Þ x~ under this formulation, we have to: (i)
assume the hidden states are available to the agent and (ii) solve
Equation 18 for the value-function. Solving for the value-function
is a non-trivial problem and usually involves backwards induction
or some approximation scheme like reinforcement-learning [4–6].
The free-energy formulation circumvents these problems by
prescribing the policy in terms of free-energy, which encodes
optimal control (Equation 6)"
  \end{itemize}

>>>>

The dmPFC subserves predominantly ambiguous amodal processes
across time, space, and content domains in
sensory-independent top-down pathways.
This DMN node has been described as a “mental sketchpad” (Goldman-Rakic, 1996),
potentially implicated in de-novo generation and binding
of meaning representations instructed by stored semantics and memories.
%
The dmPFC may thus enable inference, representation, and assessment
of one's own and other individuals' action and thoughts.
Patients with neurological lesions in the prefrontal cortex
are known to adapt to novel situations and stimuli with difficulty
(Stuss, D. T. \& Benson, D. F. The Frontal Lobes, Raven, New York, 1986).
Specifically, neural activity in human dmPFC
reflected expectations about other peoples actions and errors thereof
(Suzuki et al., 2012; van Schie et al., 2004).
dmPFC activity indeed explained the proficiency decline
of inferring other peoples thoughts in the elderly (Moran et al., 2012).
Some dmPFC neurons in macaque monkeys exhibited a preference
for processing others', rather than own, behavior
with fine-grained adjustment of contextual circumstances (Yoshida et al., 2011).
Also the topographically neighboring dorsal anterior cingulate cortex
has been linked to computing values and efforts of
persisting a behavioral plan versus switching the
environmental context in several lesion studies (Rushworth NN 2016).
%
Such highly abstract neural computations necessarily rely on the
generation of probabilistic internal information drawing from
episodic memory retrieval, generative construction processes,
and explicit knowledge of the external world.
%
According to computational neuroimaging experiments,
the dmPFC activity preferentially models stimulus-value associations of
possible actions that are not actually executed,
whereas the vmPFC activity models value in the environment that leads performed actions
{Nicolle, 2012 \#3750}.

@@

where the expectation is over all possible trajectories successor
trajectories of state $s_t$.
Given that each $\pi$ correspondence to an exhaustive description of
behavioral patterns, $Q^{\pi}(s_t,a_t)$ corresponds to the epistemic value of
the set of actions given states in finite time.

By construction, in general, maximizing the expectation in  \eqref{eq:qval} is intractable due to
the enormous number of possible trajectories to consider.
(If the Bellman-based computational ideas have a corresponding
neurobiological implementation, this provides a framework of how
optimal behavior could be computable in the human brain.)
One popular simplification that removes this issue is to only consider the sub-family of
\textit{deterministic policies}, which map each state to the best single best action which
should be taken at thin state, and therefore define functions from states to actions.

@@

\paragraph{Free Energy Principle.}
\begin{itemize}
  \item Ortega et al. 2013 \cite{ortega2013thermodynamics} ``Thermodynamics as a theory of decision-making with information-processing costs''
\item Friston's critic of RL \cite{fristonAIorRL} (he proposes "active inference''): "This equation (equation 18) comes from the theory of dynamic programming,
pioneered in the 1950s by Richard Bellman and colleagues [2]. To
optimise control a~pð Þ x~ under this formulation, we have to: (i)
assume the hidden states are available to the agent and (ii) solve
Equation 18 for the value-function. Solving for the value-function
is a non-trivial problem and usually involves backwards induction
or some approximation scheme like reinforcement-learning [4–6].
The free-energy formulation circumvents these problems by
prescribing the policy in terms of free-energy, which encodes
optimal control (Equation 6)"
\item It is difficult to compute the free energy principle
in practice, which makes it physiologically less plausible.
  \end{itemize}
